Emulation of high multiplicity matrix elements for electron-positron and hadron-hadron colliders

### 2022-10-13 ml.tex
wrote: ml intro

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Machine learning}
\label{chapter:ml}
\section{Introduction to machine learning}
\subsection{Supervised and unsupervised learning}
\section{Neural networks}
\subsection{Perceptron}
\subsection{Densely connected neural networks}
\subsection{Optimisation of weights}
\subsection{ML applications in HEP}
\end{document}

>>>>

\documentclass[main.tex]{subfiles}
\begin{document}
\chapter{Modern machine learning algorithms and their uses}
\label{chapter:ml}
\section{Rise of machine learning in HEP}
    In high energy physics where analyses
    depend on making use of large computing clusters
    and international grid efforts, advancements
    in technology and more efficient methods are
    required to keep up with the needs of the
    experimental and theoretical community.
    However, the computational resources needed
    are at risk of outpacing the growth in these research
    areas \cite{Bothmann:2022thx}. Therefore new techniques and
    algorithms from the machine learning community
    have been adopted to augment the ongoing efforts
    in the high energy physics community. Machine
    learning is a category of artificial intelligence
    concerned with the design of algorithms that
    automates learning from data.
    By now machine learning is utilised throughout
    every stage of particle physics analysis, from collection
    of data from experiments all the way to novel
    applications on the theory side.

    Around the turn of the last decade the culmination
    in advancement in hardware and techniques led to an
    explosion in image classification research in the
    machine learning community, specifically in the
    utilisation of very large neural networks. The accuracy
    of these deep neural networks greatly outperformed the
    previous state-of-the-art \cite{NIPS2012_c399862d,https://doi.org/10.48550/arxiv.1409.0575} and
    lead to the widespread use of neural networks for a
    wide array of tasks \cite{Schmidhuber_2015}.
    This led to the rise of programming
    frameworks such as scikit-learn \cite{scikit-learn},
    TensorFlow \cite{https://doi.org/10.48550/arxiv.1603.04467},
    PyTorch \cite{https://doi.org/10.48550/arxiv.1912.01703},
    and XGBoost \cite{Chen:2016:XST:2939672.2939785}
    to name a few which implement neural networks, alongside
    other mainstay algorithms such as decisions trees
    and support vector machines.
    These algorithms are now extremely commonplace
    in many domains, including in particle physics.

    Collisions at the LHC can produce hundreds of
    particles with complex final-state configurations
    leading to the design of the ATLAS and CMS experiments
    to contain of the order 100 million detection elements
    in an attempt to disentangle these events. With these
    large arrays of detection elements there is an enormouse
    amount of data harvested from each collision, terabytes
    per second, which presents a challenge in the form of
    data collection.
    Already at this stage, boosted decision trees
    have been used to enhance triggers \cite{CMS:2020cmk} to accept or
    reject data from being saved to disk. At the LHCb
    experiment 70\% of all data retained have had
    some involvement with machine learning algorithms \cite{LHCb:2014set}.

    On the theory side of things, there have been numerous
    triumphs in the application of machine learning.
    Perhaps the most well-known is the fitting
    of PDFs performed by the NNPDF collaboration \cite{NNPDF:2021njg}.
    This area is rapidly growing with many new
    ideas being explored at each stage of the event
    generation process.
    A living review that is actively being updated
    archiving advancements in the field is available
    at \cite{Feickert:2021ajf}.
    For more traditional reviews, see for instance
    \cite{Guest:2018yhq,Radovic:2018dip,Butter:2022rso}.

    In this chapter we give an overview of neural
    networks: how they are constructed, how
    they are trained, and how they are deployed.
    We will then proceed to discuss their
    applicability as surrogate models for matrix
    elements.

\section{Neural networks}
\subsection{Perceptron}
\subsection{Densely connected neural networks}
\subsection{Optimisation of weights}
\section{Neural networks as matrix element surrogate models}
\end{document}
